---
title: "Casein kinase substrate phosphorylations"
author: "Bing Shui"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Library Preparation
```{r}
suppressMessages(
  c(library(tidyverse),
    library(RColorBrewer),
    library(gplots),
    library(knitr),
    library(mosaic))
)
```

# Objective
To investigate the phosphorylation status of members of the casein kinase families and select substrates unique to Csnk2a1 for WB validation.

Casein kinase family members examined here:
  - Csnk1a1 (CK1A)
  - Csnk1d (CK1D)
  - Csnk1e (CK1E)
  - Csnk2a1 (CK2A1)
  - Csnk2a2 (CK2A2)
  - Csnk2b (CK2B)
  
Unscaled and scaled (to total protein change) phosphorylation status will be examined.

# Load datasets
```{r}
# Load kinase-substrate list from Doug Brubaker
k.s.list <-read_tsv("../kinsubset_prot_acc_mouse_20170418.gmx.txt")
k.s.list <- k.s.list[-1,]
## Select only casein kinases
k.s.list <- k.s.list %>% select(c("CK1A", "CK1D", "CK1E", "CK2A1", "CK2A2", "CK2B"))
keep_rows <- rowSums(!is.na(k.s.list))>0
k.s.list <- k.s.list[keep_rows,]

# get a list of all kinase substrates
k.s.list.all <-read_tsv("../kinsubset_prot_acc_mouse_20170418.gmx.txt")
k.s.list.all <- k.s.list.all[-1,]
substrate_list <- unlist(as.data.frame(k.s.list.all)) 
substrate_list <- substrate_list[!is.na(substrate_list)]
substrate_freq <- as_tibble(table(substrate_list))
colnames(substrate_freq)[2] <- "Num_targeting_kinase"

# Load total proteomics dataset
total_prot <- read_csv("crcMS_diff.csv")
total_prot_quant <- read_csv("2017-03-19_Haigis_5v5_Pro.csv")

# Load phosphoproteomics dataset
phos_prot <- read_csv("crcMS_phospho_diff.csv")
phos_prot$site_id <- paste(phos_prot$`Protein Id`, phos_prot$`Site Position`, sep = "_")
phos_prot$name_id <- paste(phos_prot$gene_symbol, phos_prot$`Site Position`, sep = "_")
phos_prot_quant <- read_csv("2017-03-19_Haigis_5v5_Phos.csv")
phos_prot_quant$site_id <- paste(phos_prot_quant$`Protein Id`, phos_prot_quant$`Site Position`, sep = "_")
phos_prot_quant$name_id <- paste(phos_prot_quant$gene_symbol, phos_prot_quant$`Site Position`, sep = "_")
```

# Analysis
## Set up function
```{r}
phospho_target_info <- function(kinase, all_height, sig_height) {
  # all target info
  target <- k.s.list %>% select(kinase) %>% na.omit() %>% unlist()
  target_phospho <- phos_prot %>% filter(site_id %in% target)
  target_phospho <- inner_join(target_phospho, total_prot[,c(2,7)], by = c("Protein Id" = "Protein Id"))
  target_phospho_quant <- phos_prot_quant %>% filter(site_id %in% target_phospho$site_id)
  colnames(target_phospho)[c(10,13)] <- c("Phospho_LFC", "Protein_LFC")
  target_phospho$scaled_phospho_LFC <- target_phospho$Phospho_LFC - target_phospho$Protein_LFC
  target_phospho <- inner_join(target_phospho, substrate_freq, by = c("site_id" = "substrate_list"))
  if(dim(target_phospho)[1]> 0) {
    filename <- paste("CK_Target/", kinase, "_target_phosphorylations.csv", sep = "")
    write_csv(target_phospho, filename)
  }
 
  # Significant target phosphosite info (p < 0.05)
  sig_target_phospho <- target_phospho %>% filter(p_values <= 0.05)
  sig_target_phospho_quant <- target_phospho_quant %>% filter(site_id %in% sig_target_phospho$site_id)
  print(paste("Number of significantly changed target phosphorylations of ", kinase, ": ", dim(sig_target_phospho)[1], sep = ""))
  
  if (dim(sig_target_phospho)[1] > 0) {
    filename <- paste("CK_Target/", kinase, "_sig_target_phosphorylations.csv", sep = "")
    write_csv(sig_target_phospho, filename)
  }
  # Heatmap of all targets (unscaled and scaled)
  my_palette <- colorRampPalette(c("blue", "white", "red"))(100)
  if (dim(target_phospho)[1] > 1) {
    # LFC heatmap
    heatmap_matrix <- as.matrix(target_phospho[,c(10,14)])
    rownames(heatmap_matrix) <- target_phospho$name_id
    figure_name <- paste("CK_Target/", kinase, "_target_phosphorylations_LFC.png", sep = "")
    png(figure_name,
        width = 600,
        height = all_height,
        res = 100,
        pointsize = 8)
    par(cex.main=1.1)
    heatmap.2(heatmap_matrix,
              main = paste(kinase, "target phosphorylations LFC"),
              density.info = "none",
              key = TRUE,
              lwid = c(3,7),
              lhei = c(1,7),
              col=my_palette,
              margins = c(15,10),
              symbreaks = TRUE,
              breaks=seq(-2,2,0.04),
              trace = "none",
              dendrogram = "row",
              labRow = rownames(heatmap_matrix),
              cexCol = 1.5,
              Colv = "NA")
    dev.off()
    
    # Quantification heatmap
    heatmap_matrix <- as.matrix(target_phospho_quant[, c(6:15)])
    heatmap_matrix <- t(apply(heatmap_matrix, 1, zscore))
    rownames(heatmap_matrix) <- target_phospho_quant$name_id
    figure_name <- paste("CK_Target/", kinase, "_target_phosphorylations_quant.png", sep = "")
    png(figure_name,
        width = 600,
        height = all_height,
        res = 100,
        pointsize = 8)
    par(cex.main=1.1)
    heatmap.2(heatmap_matrix,
              main = paste(kinase, "target phosphorylations Quant"),
              density.info = "none",
              key = TRUE,
              lwid = c(3,7),
              lhei = c(1,7),
              col=my_palette,
              margins = c(15,10),
              symbreaks = TRUE,
              trace = "none",
              dendrogram = "row",
              labRow = rownames(heatmap_matrix),
              cexCol = 1.5,
              Colv = "NA")
    dev.off()
  }
  
  # Heatmap of significant targets (unscaled and scaled)
  if (dim(sig_target_phospho)[1] > 1) {
    # Sig LFC heatmap
    heatmap_matrix <- as.matrix(sig_target_phospho[,c(10,14)])
    rownames(heatmap_matrix) <- sig_target_phospho$name_id
    figure_name <- paste("CK_Target/", kinase, "_sig_target_phosphorylations_LFC.png", sep = "")
    png(figure_name,
        width = 600,
        height = sig_height,
        res = 100,
        pointsize = 8)
    par(cex.main=1.1)
    heatmap.2(heatmap_matrix,
              main = paste(kinase, "sig target phosphorylations LFC"),
              density.info = "none",
              key = TRUE,
              lwid = c(3,7),
              lhei = c(1,7),
              col=my_palette,
              margins = c(15,10),
              symbreaks = TRUE,
              breaks=seq(-2,2,0.04),
              trace = "none",
              dendrogram = "row",
              labRow = rownames(heatmap_matrix),
              cexCol = 1.5,
              Colv = "NA")
    dev.off()
    
    # Sig Quantification heatmap
    heatmap_matrix <- as.matrix(sig_target_phospho_quant[, c(6:15)])
    heatmap_matrix <- t(apply(heatmap_matrix, 1, zscore))
    rownames(heatmap_matrix) <- sig_target_phospho_quant$name_id
    figure_name <- paste("CK_Target/", kinase, "_sig_target_phosphorylations_quant.png", sep = "")
    png(figure_name,
        width = 600,
        height = sig_height,
        res = 100,
        pointsize = 8)
    par(cex.main=1.1)
    heatmap.2(heatmap_matrix,
              main = paste(kinase, "sig target phosphorylations Quant"),
              density.info = "none",
              key = TRUE,
              lwid = c(3,7),
              lhei = c(1,7),
              col=my_palette,
              margins = c(15,10),
              symbreaks = TRUE,
              trace = "none",
              dendrogram = "row",
              labRow = rownames(heatmap_matrix),
              cexCol = 1.5,
              Colv = "NA")
    dev.off()
  }
}

```

## Csnk1a1 (CK1A)
```{r}
phospho_target_info("CK1A", 800, 600)
figures <- list.files("CK_Target", ".*CK1A.*\\.png$")
if (length(figures)>0) {
  paths <- paste0("CK_Target/", figures)
  include_graphics(paths)
}
```

## Csnk1d (CK1D)
```{r}
phospho_target_info("CK1D", 800, 600)
figures <- list.files("CK_Target", ".*CK1D.*\\.png$")
if (length(figures)>0) {
  paths <- paste0("CK_Target/", figures)
  include_graphics(paths)
}
```

## Csnk1e (CK1E)
```{r}
phospho_target_info("CK1E", 800, 600)
figures <- list.files("CK_Target", ".*CK1E.*\\.png$")
if (length(figures)>0) {
  paths <- paste0("CK_Target/", figures)
  include_graphics(paths)
}
```

## Csnk2a1 (CK2A1)
```{r}
phospho_target_info("CK2A1", 1400, 800)
figures <- list.files("CK_Target", ".*CK2A1.*\\.png$")
if (length(figures)>0) {
  paths <- paste0("CK_Target/", figures)
  include_graphics(paths)
}
```

## Csnk2a2 (CK2A2)
```{r}
phospho_target_info("CK2A2", 600, 800)
figures <- list.files("CK_Target", ".*CK2A2.*\\.png$")
if (length(figures)>0) {
  paths <- paste0("CK_Target/", figures)
  include_graphics(paths)
}
```

## Csnk2b (CK2B)
```{r}
phospho_target_info("CK2B", 600, 800)
figures <- list.files("CK_Target", ".*CK2B.*\\.png$")
if (length(figures)>0) {
  paths <- paste0("CK_Target/", figures)
  include_graphics(paths)
}
```

# Session Info
```{r}
sessionInfo()
```